---
title: "apple non-seasonal"
output: pdf_document
date: "2024-04-30"
---

```{r}
library(stats)
library(tseries)
library(tidyverse)
library(TSA)
library(ggplot2)
library(rugarch)
library(forecast)
library(dplyr)
```

```{r}
data <- read.csv("/Users/shivanipatel/Downloads/AAPL.csv")

head(data)

```

```{r}
data$Date <- as.Date(data$Date, format="%Y-%m-%d")
```

```{r}
summary(data)
```

```{r}
# Plotting the closing prices over time
ggplot(data, aes(x = Date, y = Close)) +
  geom_line() +
  labs(title = "S&P 500 Close Prices Over Time", x = "Date", y = "Close Price")
```

```{r}
sum(is.na(data$Close))
```
```{r}
data <- select(data, Date, Close)
```

```{r}
data <- data[sample(nrow(data)), ]

n_rows <- nrow(data)
train_rows <- round(n_rows * 0.80)

train_data <- data[1:train_rows, ]
test_data <- data[(train_rows + 1):n_rows, ]

tail(train_data)
```


```{r}
tail(train_data)
```

```{r}
data1 <- train_data$Close
```


```{r}
acf(data1, lag.max = 150, main='ACF of Original data')
```

```{r}
pacf(data1, lag.max = 150, main='PACF of Original data')
```

```{r}
eacf(data1)
```

```{r}
adf_test = adf.test(data1)
print(adf_test)
print(" the TS is non-stationary")
```

The Augmented Dickey-Fuller test result shows a Dickey-Fuller statistic of -20.086 with a p-value of 0.01, suggesting the time series data is statistically significant for stationarity, contradicting the note that the time series is "non-stationary". This could indicate a typo or misinterpretation; a p-value of 0.01 generally supports the rejection of the null hypothesis that the series has a unit root (i.e., it is non-stationary).

```{r}
close_diff <- diff(data1)
plot(close_diff, type = 'l', xlab = 'Date', ylab = 'Value', main = 'Line graph for log-differencing')
```

The line graph shows log-differenced data, typically used to stabilize variance and achieve stationarity in time series, like stock prices. The plot reveals substantial volatility and potential outliers, indicating active market dynamics. This transformation is essential for preparing the data for further analysis, such as identifying suitable ARIMA model parameters using autocorrelation functions. The data appears noisy, suggesting the potential need for further smoothing or differencing.

```{r}
adf_test = adf.test(close_diff)
print(adf_test)
```


The output from the Augmented Dickey-Fuller (ADF) test shows a Dickey-Fuller statistic of -34.736 with a p-value of 0.01, suggesting strong evidence against the null hypothesis of non-stationarity. This indicates that the log-differenced data is stationary, meaning it has a constant mean and variance over time, making it suitable for further time series analysis and forecasting.

## ACF & PACF PLOTS

```{r}
acf(close_diff, main='ACF of log-differenced data')
```

```{r}
pacf(close_diff, main='PACF of log-differenced data')
```

```{r}
eacf(close_diff)
```

## MODELING

## ARIMA MODEL

```{r}
plot_arima_resid <- function(model) {
  resid <- resid(model)
  resid_ts <- ts(resid, start=c(2013, 311), end=c(2020,300), frequency=365)
  p <- model$arma[1]
  q <- model$arma[2]
  plot(resid_ts, xlab = "Year", ylab = 'Residuals', main=paste('Residual Plot',"(", p, ", 1, ", q, ")"))
  acf(resid_ts[1:length(resid_ts)],lag.max=150, main=paste('ACF of Residual',"(", p, ", 1, ", q, ")"))
  pacf(resid_ts[1:length(resid_ts)],lag.max=150, main=paste('PACF of Residual',"(", p, ", 1, ", q, ")"))
}
```

```{r}
my_data <- data.frame(model = character(),
                    AIC = numeric(),
                    BIC = numeric(),
                    Shapiro = round(numeric(),3),
                    Ljung = round(numeric(),3))
new_row <- data.frame(model='',AIC=0,BIC=0,Shapiro=0,Ljung=0)
```

```{r}
plot_arima_residuals <- function(arima_model, my_data) {
  resid <- residuals(arima_model)
  resid_ts <- ts(resid, start=c(2013, 311), end=c(2020,300), frequency=365)
  p <- arima_model$arma[1]
  q <- arima_model$arma[2]
  order <- paste("(", p, ", 1, ", q, ")", sep = "")
  plot_arima_resid(arima_model)
  qqnorm(resid_ts, main=paste('Residuals plot',order))
  qqline(resid_ts)

  shap=shapiro.test(resid_ts)

  ljung=Box.test(resid, lag = 20, type = "Ljung-Box")

  cat("AIC:", AIC(arima_model),"\n")
  cat("BIC:", BIC(arima_model),"\n")
  model_name <- paste("ARIMA", paste(order, collapse = ','), sep = " ")
  if (!model_name %in% my_data$model) {
    new_row <- data.frame(model = model_name,
                          AIC = AIC(arima_model),
                          BIC = BIC(arima_model),
                          Shapiro = round(shap$p.value, 3),
                          Ljung = round(ljung$p.value, 3))

    my_data <- rbind(my_data, new_row)
  }
}
```

```{r}
arima011 <- arima(x=close_diff, order=c(0,1,1))
arima110 <- arima(x=close_diff, order=c(1,1,0))
arima111 <- arima(x=close_diff, order=c(1,1,1))
arima210 <- arima(x=close_diff, order=c(2,1,0))
arima211 <- arima(x=close_diff, order=c(2,1,1))
arima012 <- arima(x=close_diff, order=c(0,1,2))
arima112 <- arima(x=close_diff, order=c(1,1,2))
arima212 <- arima(x=close_diff, order=c(2,1,2))
arima013 <- arima(x=close_diff, order=c(0,1,3))
arima113 <- arima(x=close_diff, order=c(1,1,3))
arima213 <- arima(x=close_diff, order=c(2,1,3))
arima310 <- arima(x=close_diff, order=c(3,1,0))
arima311 <- arima(x=close_diff, order=c(3,1,1))
arima312 <- arima(x=close_diff, order=c(3,1,2))
arima313 <- arima(x=close_diff, order=c(3,1,3))
```

```{r}
my_data <- plot_arima_residuals(arima011, my_data)
```


```{r}
my_data <- plot_arima_residuals(arima011, my_data)
my_data <- plot_arima_residuals(arima110, my_data)
my_data <- plot_arima_residuals(arima111, my_data)
my_data <- plot_arima_residuals(arima210, my_data)
my_data <- plot_arima_residuals(arima211, my_data)
my_data <- plot_arima_residuals(arima012, my_data)
my_data <- plot_arima_residuals(arima112, my_data)
my_data <- plot_arima_residuals(arima212, my_data)
my_data <- plot_arima_residuals(arima013, my_data)
my_data <- plot_arima_residuals(arima113, my_data)
my_data <- plot_arima_residuals(arima213, my_data)
my_data <- plot_arima_residuals(arima310, my_data)
my_data <- plot_arima_residuals(arima311, my_data)
my_data <- plot_arima_residuals(arima313, my_data)
my_data <- plot_arima_residuals(arima312, my_data)
```
## ARIMA Model Results

```{r}
my_data
```

```{r}
BIC_values <- c(BIC(arima011), BIC(arima110), BIC(arima111),
                BIC(arima210), BIC(arima211), BIC(arima012), BIC(arima112), BIC(arima212),
                BIC(arima310), BIC(arima311), BIC(arima312), BIC(arima013), BIC(arima113), BIC(arima213),   
                BIC(arima313))
BIC_values
```

```{r}
min_BIC <- min(BIC_values)
min_BIC
```

Finding the minimum BIC Values to find the best fit, and turns out arima113 is the best fit among all.


```{r}
arima113

```
# RESIDUAL ANALYSIS

```{r}
residuals <- residuals(arima113)
residuals
```

```{r}
acf(residuals, main = 'ACF of Residuals')
pacf(residuals, main = 'PACF of Residuals')
# Ljung-Box test for residual autocorrelation
tsdiag(arima113)
```

## ACF & PACF PLOTS

```{r}
predicted_values <- predict(arima113, n.ahead = 1981)
print(predicted_values)
test_data$Close
original_forecasts <- diffinv(predicted_values$pred, lag = 1, differences = 1, xi = 39.84)
original_forecasts <- original_forecasts[original_forecasts != 0]
#original_forecasts <- original_forecasts[-1]
original_forecasts
```


```{r}

original_forecasts
```


```{r}
length(test_data$Date)
length(original_forecasts)
```


```{r}
# Length Checks
length(test_data$Date)
length(original_forecasts)

# Plotting Section (Updated)
plot(test_data$Date, test_data$Close, main = 'Original Data vs Forecast for 2024',
     type = 'l', col = 'black', xlab = 'Date', ylab = 'Close',
     xlim = c(min(test_data$Date), max(test_data$Date)), # Adjust x-axis range if needed
     ylim = c(min(c(test_data$Close, original_forecasts)),  # Adjust y-axis range if needed
              max(c(test_data$Close, original_forecasts))))
lines(test_data$Date, original_forecasts, col = 'red', type = 'l', lwd = 2)
```


```{r}
# Define GARCH model specification
garch_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(2, 2)),
                         mean.model = list(armaOrder = c(3, 4), include.mean = TRUE),
                         distribution.model = "sstd")

return_data <- diff(log(data$Close))[-1]

# Fit the GARCH model
garch_fit <- ugarchfit(spec = garch_spec, data = return_data, solver = 'hybrid')  # ensure 'return_data' is defined
summary(garch_fit)

```


```{r}
for (i in 1:12) {
  plot(garch_fit, which = i)
}
```

```{r}
plot(garch_fit, which = 12)

```

```{r}
garch_fit
```


```{r}
resid_norm = residuals(garch_fit)
qqnorm(resid_norm, main='QQ plot of residuals from GARCH ')
qqline(resid_norm)
acf(resid_norm, lag.max = 60, main='ACF of residual')
pacf(resid_norm, lag.max = 60, main='PACF of residual')
```

```{r}
initial_term <- train_data$Close[length(train_data$Close)]
# Forecast next 30 days
forecasted <- ugarchforecast(garch_fit, n.ahead = 30, startMethod = "sample", initial.level = initial_term)
# Extract the forecasted values
forecast_values <- fitted(forecasted)
```

```{r}
predicted_diff <- cumsum(c(0, forecast_values))
predicted_log <- log(initial_term) + predicted_diff
predicted_close <- exp(predicted_log)
print(predicted_close[-1])

```

```{r}
plot(test_data$Date, test_data$Close, main = 'Original Data vs Forecast for 30 days', type = 'l', col = 'red', xlab = 'Date', ylab = 'Employees')

```

```{r}
conditional_variance <- garch_fit@fit$sigma^2
std_resid <- resid_norm / sqrt(conditional_variance)
plot(std_resid, type = "l", ylab = "Standardized Residuals")
```

```{r}
Box.test(resid_norm, lag = 20, type = "Ljung-Box")
```

```{r}
return = abs(diff(log(train_data$Close)))
```

```{r}
acf(return, lag.max = 100, main='ACF of return')
pacf(return, lag.max = 100, main='PACF of return')
eacf(return)
garch(x=return, grad='numerical', trace=FALSE)
garch_spec <- ugarchspec(variance.model = list(garchOrder = c(2,2)),
                         mean.model = list(armaOrder = c(3,5)),
                         distribution.model = 'std')
```

```{r}
garch_fit <- ugarchfit(spec = garch_spec, data = return, solver = 'hybrid')
summary(garch_fit)
garch_fit
```

```{r}
resid_norm = residuals(garch_fit)
qqnorm(resid_norm, main='QQ plot of residuals from GARCH ')
qqline(resid_norm)
acf(resid_norm, lag.max = 60, main='ACF of residual')
pacf(resid_norm, lag.max = 60, main='PACF of residual')
```

```{r}
initial_term <- train_data$Close[length(train_data$Close)]
# Forecast next 10 days
forecasted <- ugarchforecast(garch_fit, n.ahead = 10, startMethod = "sample")
# Extract the forecasted values
forecast_values <- fitted(forecasted)
predicted_diff <- cumsum(c(0, forecast_values))
predicted_log <- log(initial_term) + predicted_diff
predicted_close <- exp(predicted_log)
print(predicted_close[-1])
```

```{r}
plot(test_data$Date, test_data$Close, main = 'Original Data vs Forecast for 30 days', type = 'l', col = 'red', xlab = 'Date', ylab = 'Employees')

```

```{r}
conditional_variance <- garch_fit@fit$sigma^2
std_resid <- resid_norm / sqrt(conditional_variance)
plot(std_resid, type = "l", ylab = "Standardized Residuals")

```


```{r}
Box.test(resid_norm, lag = 20, type = "Ljung-Box")

```

```{r}
return = diff(log(train_data$Close))^2
```

```{r}
garch(x=return, grad='numerical', trace=FALSE)
garch_spec <- ugarchspec(variance.model = list(garchOrder = c(2,2)),
                         mean.model = list(armaOrder = c(2,2)),
                         distribution.model = 'std')
```

```{r}
garch_fit <- ugarchfit(spec = garch_spec, data = return, solver = 'hybrid')
summary(garch_fit)
garch_fit
```

```{r}
resid_norm = residuals(garch_fit)
qqnorm(resid_norm, main='QQ plot of residuals from GARCH ')
qqline(resid_norm)
acf(resid_norm, lag.max = 60, main='ACF of residual')
pacf(resid_norm, lag.max = 60, main='PACF of residual')
```

```{r}
sfinal <- garch_spec
setfixed(sfinal) <- as.list(coef(garch_fit))
coef(garch_fit)
f2019 <- ugarchforecast(data = return, fitORspec = sfinal, n.ahead = 60)
plot(sigma(f2019))
sim <- ugarchpath(spec = sfinal, m.sim = 2, n.sim = 1*60, rseed = 123)
p <- 39.84*apply(fitted(sim), 2, 'cumsum') + 39.84
matplot(p, type = "l", lwd = 3)
```

```{r}
garch_spec <- ugarchspec(variance.model = list(model = 'gjrGARCH'),
                         mean.model = list(armaOrder = c(2,2)),
                         distribution.model = 'std')

garch_fit <- ugarchfit(spec = garch_spec, data = return, solver = 'hybrid')
summary(garch_fit)
garch_fit
```

```{r}
resid_norm = residuals(garch_fit)
qqnorm(resid_norm, main='QQ plot of residuals from GARCH ')
qqline(resid_norm)
acf(resid_norm, lag.max = 60, main='ACF of residual')
pacf(resid_norm, lag.max = 60, main='PACF of residual')
```

```{r}
sfinal <- garch_spec
setfixed(sfinal) <- as.list(coef(garch_fit))
coef(garch_fit)
f2019 <- ugarchforecast(data = return, fitORspec = sfinal, n.ahead = 25)
plot(sigma(f2019))
sim <- ugarchpath(spec = sfinal, m.sim = 2, n.sim = 1*60, rseed = 123)
p <- 39.84*apply(fitted(sim), 2, 'cumsum') + 39.84
matplot(p, type = "l", lwd = 3)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
